---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# metrdata

<!-- badges: start -->
<!-- badges: end -->

The goal of metrdata is to provide a simply interface for 
associating arbitrary metadata with data in R. These metadata can, for example, 
be used to describe the source of the data, any processing steps taken, and 
parameters used to process the data.

These metadata should be human readable so are saved alongside the data files
in sidecar files in JSON format.

Given that data files are likely to derive from other sources, functions
to inherit metadata from other objects are provided.


## Installation

You can install the development version of metrdata like so:

``` r
library(devtools)
install_github("twlee79/metrdata")
```

## Metadata storage within an R session

Metadata are stored in an R session by assigning as a attribute for a data
object. The metadata for an object are stored in the attribute named
``r metrdata::get_metrdata_attrname()`` as an S3 class with the class name 
``r metrdata::get_metrdata_classname()``.

## Metadata persistent storage

A goal of this package is to ensure that metadata of a data file are human
readable. Therefore, the metadata are stored as sidecar files 
(i.e. separate files) with the suffix ``r metrdata::get_metrdata_suffix()`` 
alongside the data file in JSON format.


## Example: Suggested metadata usage pattern

Although metrdata allows the use of arbitrary metadata, a suggest pattern is
to assign metadata under lists using top-level 'keys'. This allows metadata
from multiple steps of processing to be better isolated under separate keys.
In addition, this allows a user-friendly means of inherting metadata from
previous processing steps.

For example, consider this starting data with metadata applied:
```{r}
library(metrdata)
cars_data<-set_metadata(cars, 
  metrdata(`Original data source` = "cars dataset from R"))

metadata(cars_data)
```

And these processing functions, which replace metadata with new metadata
providing details on the processing and reference the source dataset:
```{r}
process_cars_speed<-function(dataset) {
  dataset$speed<-dataset$speed*0.9
  set_metadata(
    dataset, replace = TRUE,
    metrdata(
      `Processing of speed` = list(
        Note = "speed alteration applied",
        Amount = "reduced by 10%"
      ),
      dataset
      
    )
  )
}

process_cars_dist<-function(dataset, amount = 2) {
  # set metadata for this processing function
  .metadata_label <- "Processing of dist"
  .metadata_list <- list(
    Note = "dist alteration: multiplication applied",
    Amount = amount
  )
  .metadata_source <- dataset 
    # or list(dataset, ...) for >1 source
  
  result<-strip_metadata(dataset)
    # alternative is to replace = TRUE in set_metadata() 
  result$speed<-result$speed*2
  
  # return result dataset with added metadata
  set_metadata(
    result,
    metrdata(
      label =.metadata_label,
      .metadata_list,
      include = .metadata_source
    )
  )
}
```

The result is a easily understood metadata object with each step of processing 
outlined under a descriptive heading (the 'key'):
```{r}
derived_cars<-process_cars_speed(cars_data)
metadata(derived_cars)

derived_cars2<-process_cars_dist(derived_cars)
metadata(derived_cars2)
```

Note the pattern for the second processing function `process_cars_dist()`,
where the details of the metadata (label, metadata, sources) are recorded 
at the top of function, and applied when the function returns. This pattern
is useful for keeping the metadata description at the top of the function,
where it may be easier to maintain.

## History

v0.1.0 
  - 2024-02-05 
  - Initial release
